---
title: "fibers - C1 - FAE - no comment"
author: "Your Name"
date: "The Date"
output: 
  html_document:
    css: ../makeConfig.css
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(htmlTable)
library(dplyr)
library(readxl)
library(data.table)
library(jsonlite)

devtools::load_all(Sys.getenv("FUSION"))
devtools::load_all(Sys.getenv("ROLODEX"))
set.seed(5474)
options("max.print" = 10000)
```

# project identification
```{r}
# Get the name of the project from the database
con <- openDb()
project <- input(con, what = "project")
```

```{r}
# Set project and prefix
projectName <- project$name
projectID <- project$projectID
prefix <- rolProject(projectName)$prefix
```

```{r}
# Get the name of the cohort name from the database
cohortName <- input(con,
                    what = "cohort",
                    options = list("projectID" = projectID))$cohortName
```

# selecting and reading manifest
```{r}
# Get file with manifest. If using xlsx files, make sure that the information
# is on the first sheet
manifest <- as.data.table(inputFile())
N <- nrow(manifest)
```

```{r}
# creating sampleIDs
sampleID <- getLastSampleID(projectName, length.out = N, offset = 0)$nextID
```

# matching and checking
```{r} 
# Match column in the manifest with the mandatory fields below.
# If dates are important, please make sure that import is successful.
# Dates often get converted automatically by excel and by R!!!

# dates <- as.POSIXct(dates, tz = "UTC", format = "%d/%m/%y")

manifest$sampleID <- sampleID
manifest$projectName <- rep(projectName, N)
manifest$cohortName <- rep(cohortName, N)

# creating date for today (comment after use)
today <- format(Sys.time(), "%d-%m-%Y")
manifest$receptionDate <- rep(today, N)

# sampleMatrixTypes are predefined. Use getMatrices() to obtain a list of
# valid types. Most common are PLA, SER and URI
manifest$sampleMatrixType <- rep("FAE", N)

# sample volume must be in microliter without unit (eg. 500)
manifest$sampleVolume <- rep(NA, N)

# tubeLabel MUST correspond to the labels actually on the tube so that 
# locating samples is possible
manifest$tubeLabel <- cleanNames(manifest$tubeLabel)

# sourceID corresponds to the patientID or the source in a multi time point
# experimental design.
# IN ORDER TO UNIQUELY MATCH A SAMPLE TO ITS METADATA THE CONCATENATION OF
# SOURCEID - SAMPLETIMEPOINT - SAMPLEALIQUOTS MUST BE UNIQUE.
manifest$sourceID <- cleanNames(manifest$sourceID)
manifest$sampleTimePoint <- paste0(cleanNames(manifest$sampleTimePoint),
                                   cleanNames(manifest$Visit))
manifest$sampleAliquots <- rep(1, N)

# Provide information about the boxes
# Please use:
# 9x9
# 10x10
# 9x9 (Lettered)
# 10x10 (Lettered) 
# as boxTypes.
manifest$boxName <- rep(NA, N)
manifest$boxType <- rep(NA, N)

# the position MUST be lettered. If boxes uses numeric position, then the 
# position must be converted.
manifest$tubePosition <- rep(NA, N)
manifest$tubeRow <- rep(NA, N) #posToRC(manifest$tubePosition)$row
manifest$tubeCol <- rep(NA, N) #posToRC(manifest$tubePosition)$col

# additional information
manifest$collectionTube <- rep(NA, N)
manifest$comments <- manifest$Food

# DON'T ALTER THIS LINE (COMPATIBILITY WITH OLDER FORMATS)
manifest$UUID <- paste0(manifest$sampleID, "_", manifest$sourceID)
```

# check for duplicated
```{r}
uniqueKey <- paste0(manifest$sourceID, manifest$sampleTimePoint, manifest$sampleAliquots, manifest$sampleMatrixType)
fi <- duplicated(uniqueKey)
table(fi)

if (any(fi)) {
  fi <- dduplicated(paste0(manifest$sourceID, manifest$sampleTimePoint, manifest$sampleAliquots, manifest$sampleMatrixType))
  
  # we print out which samples are duplicated in a nice graphical way
  dups <- manifest[fi,c("sourceID", "sampleTimePoint", "sampleAliquots", "sampleMatrixType")]
  setDT(dups)
  dups[order(sourceID, sampleTimePoint, sampleAliquots, sampleMatrixType)]
}

manifest$sampleAliquots <- as.numeric(makeUnique(uniqueKey, flagOnly = TRUE))
manifest$tubeLabel <- paste0(manifest$tubeLabel, "-",
                             as.numeric(makeUnique(manifest$tubeLabel, flagOnly = TRUE)))
```

# add complementary information into the comment column 
```{r}
# # use this for external projects only
# for (i in 1:nrow(manifest)) {
#   manifest$comments[i] <- toJSON(manifestOriginal[i,], unbox = TRUE)
# }
```

# creation of the sample description list
```{r}
manifest %>% select(sampleID,
                    projectName,
                    cohortName,
                    receptionDate,
                    sampleMatrixType,
                    sampleVolume,
                    tubeLabel,
                    sourceID,
                    sampleTimePoint,
                    sampleAliquots,
                    boxName,
                    boxType,
                    tubePosition,
                    tubeRow,
                    tubeCol,
                    collectionTube,
                    comments,
                    UUID) -> sampleDescriptionList
```

# checking a few things
```{r}
checkSamples(sampleDescriptionList)

```

# storing the list in the repo (legacy)
```{r}
# we still write a sample description list file for legacy purpose
# we add ignore so it is not imported
# DON'T ALTER THE NAME OF THE FILE
txt = "FAE" # if you really need to identify multiple lists, please use this.

write.table(sampleDescriptionList, 
            file = paste0("sampleDescriptionList", cleanNames(txt), ".ignore.tsv"), sep = "\t", dec = ".", row.names = FALSE)
```

# inserting into the db
```{r}
# uncomment this line to insert the samples into rolodex. 
# BEFORE DOING SO, MAKE SURE YOU HAVE PULLED THE ROLODEX-SAMPLES REPO!!!
# DON'T FORGET TO COMMIT AND PUSH ONCE INSERTED!!!
# insertSamples(sampleDescriptionList)

# and disconnecting DB
dbDisconnect(con)
```

# RLL forgot some samples (she added to the sampleDescList)
```{r}
tt <- read.table("originalFiles/sampleDescriptionListfae.ignore.tsv", header = TRUE, sep = "\t")
tt <- tt[25:28,1:18]
checkSamples(tt)
tt$boxName <- cleanNames(tt$boxName)
# insertSamples(tt)
```

NO NEED TO MODIFY THE FOLDERS AS NO RUN HAS BEEN DONE WITH THIS BATCH YET

now we need to modify plates
```{r}
lastIdx <- dbGetQuery(con, "select idx from plates order by idx desc limit 1")
idx <- seq_len(4) + as.numeric(lastIdx)

pos <- c("C4", "C5", "C6", "C7")
```

FIB01142|H5|FIBp024|fibers|C1|URI|48977|FIBp024
```{r}
projectName <- "fibers"
cohortName <- "C1"
sampleMatrixType <- "FAE"
plateID <- "FIBp045"
qry <- paste0("insert into plates ('sampleID', 
              'wellPos', 
              'plateID', 
              'projectName', 
              'cohortName', 
              'sampleMatrixType', 
              'idx',
              'plateOriginal') values
              ('", tt$sampleID[1], "',
              '", pos[1], "',
              '", plateID, "',
              '", projectName,"',
              '", cohortName,"',
              '", sampleMatrixType,"',
              '", idx[1],"',
              '", plateID,"')")
# dbExecute(con, qry)

qry <- paste0("insert into plates ('sampleID', 
              'wellPos', 
              'plateID', 
              'projectName', 
              'cohortName', 
              'sampleMatrixType', 
              'idx',
              'plateOriginal') values
              ('", tt$sampleID[2], "',
              '", pos[2], "',
              '", plateID, "',
              '", projectName,"',
              '", cohortName,"',
              '", sampleMatrixType,"',
              '", idx[2],"',
              '", plateID,"')")
# dbExecute(con, qry)

qry <- paste0("insert into plates ('sampleID', 
              'wellPos', 
              'plateID', 
              'projectName', 
              'cohortName', 
              'sampleMatrixType', 
              'idx',
              'plateOriginal') values
              ('", tt$sampleID[3], "',
              '", pos[3], "',
              '", plateID, "',
              '", projectName,"',
              '", cohortName,"',
              '", sampleMatrixType,"',
              '", idx[3],"',
              '", plateID,"')")
# dbExecute(con, qry)

qry <- paste0("insert into plates ('sampleID', 
              'wellPos', 
              'plateID', 
              'projectName', 
              'cohortName', 
              'sampleMatrixType', 
              'idx',
              'plateOriginal') values
              ('", tt$sampleID[4], "',
              '", pos[4], "',
              '", plateID, "',
              '", projectName,"',
              '", cohortName,"',
              '", sampleMatrixType,"',
              '", idx[4],"',
              '", plateID,"')")
# dbExecute(con, qry)
```

```{r}
dbGetQuery(con, "select * from samples where tubeLabel = 'fib329-f1-t1-1'")
dbExecute(con, "update samples set tubeLabel = 'fib329-f1-t0-1' where tubeLabel = 'fib329-f1-t1-1';")
dbExecute(con, "update samples set sampleTimePoint = 't0f1' where tubeLabel = 'fib329-f1-t0-1';")
dbGetQuery(con, "select * from samples where tubeLabel = 'fib329-f1-t0-1'")


dbGetQuery(con, "select * from samples where tubeLabel = 'fib439-f1-t1-1'")
dbExecute(con, "update samples set tubeLabel = 'fib439-f1-t0-1' where tubeLabel = 'fib439-f1-t1-1';")
dbExecute(con, "update samples set sampleTimePoint = 't1f1' where tubeLabel = 'fib439-f1-t0-1';")
dbGetQuery(con, "select * from samples where tubeLabel = 'fib439-f1-t0-1'")


dbGetQuery(con, "select * from samples where tubeLabel = 'fib439-f3-t1-1'")
dbExecute(con, "update samples set tubeLabel = 'fib439-f3-t0-1' where tubeLabel = 'fib439-f3-t1-1';")
dbExecute(con, "update samples set sampleTimePoint = 't0f3' where tubeLabel = 'fib439-f3-t0-1';")
dbGetQuery(con, "select * from samples where tubeLabel = 'fib439-f3-t0-1'")
```



