---
title: "Fiber Urine - NMR analysis"
author: "RM"
date: "May 13th Tue 2025"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(htmlTable)
library(dplyr)
library(readxl)
library(data.table)
library(fusion)
library(nmr.spectra.processing)
library(mva.plots)
library(ggplot2)

```

```{r}

an<-local(get(load("~/OneDrive - Murdoch University/datasets/fibers/C1/dataElements/fibers_C1_ANN.daE")))

an<-an@obsDescr[[1]]
da <- local(get(load("~/Downloads/fibers_C1_URI_FIBr02_noesygppr1d@PROF_URINE_NOESY.daE")))

spec<-da@.Data
ppm<-as.numeric(da@varName)
spec_an<-da@obsDescr[[1]]
colnames(spec)<-ppm
rownames(spec)<-spec_an$sampleID

```


*assume the LTR checks are done*

```{r}
spec<-spec[which(spec_an$sampleType=="sample"),]
spec_an<-spec_an[which(spec_an$sampleType=="sample"),]
length(which(an$sampleID %in% spec_an$sampleID))
cat("\n", "we have", length(which(an$sampleID %in% spec_an$sampleID)), " samples in common\n",
    "and ", length(which(!spec_an$sampleID %in% an$sampleID)), " samples in the experimental data that are not in metadata \n")
spec<-spec[which(spec_an$sampleID %in% an$sampleID),]
spec_an<-spec_an[which(spec_an$sampleID %in% an$sampleID),]
an<-an[match(spec_an$sampleID,an$sampleID),]

```

### NMR spectra post processing

### Step 1 Spectra calibration

Lets check the alanine doublets as it is

```{r}
matspec(spec,ppm,roi = c(1.465,1.52),interactive = F)
```


```{r}
X_cal<-calibrateSpectra(ppm,spec,ref = c("alanine"),cshift = 1.48,maxShift = 1.5,rOref = c(1.47,1.51))
colnames(X_cal)<-ppm
```

Lets see after calibration

```{r}
matspec(X_cal,ppm,roi = c(1.465,1.52),interactive = F)

```

### Step 2 Remove the regions of the spectra that are not of interest

```{r}
idx<-c(which(ppm<=0.5),which(ppm>=4.7 & ppm<=5.0),which(ppm>=9.5))

X<-X_cal[,-idx]
ppm<-ppm[-idx]

rm(idx,X_cal)
```

### Step 3 Baseline correction
```{r}
X<-baselineCorrection(X)

matspec(X,ppm,roi = c(1.465,1.52),interactive = F)

```

### Step 4 Normalization
```{r}
X_pqn<-normalize(X)
colnames(X_pqn)<-ppm

matspec(X_pqn,ppm,roi = c(1.465,1.52),interactive = F)

```


```{r}
# da<-local(get(load("~/Downloads/fibers_C1_URI_FIBr02_brxsm.daE")))
# sm<-data.frame(apply(da@.Data,2,as.numeric))
# sm_an<-da@obsDescr[[1]]
```


```{r}
# table(an$type)
#              Apple  Baseline diet only     Elephant Garlic 
#                  53                  25                  50 
#     Globe Artichoke Jerusalem Artichoke              Orange 
#                  50                  44                  47 
#              Tomato          Watermelon               Yacon 
#                  45                  50                  49 
```


# Analysis

## PCA

```{r}
mod<-PCA(X_pqn)

plotScores(
  mod,
  optns = list(
    color = an$type,
    shape = an$timePoint,
    discretePalette = c(
      "Apple" = "#FF0000",
      "Baseline diet only" = "#00FF00",
      "Elephant Garlic" = "#0000FF",
      "Globe Artichoke" = "#FFFF00",
      "Jerusalem Artichoke" = "#FF00FF",
      "Orange" = "#00FFFF",
      "Tomato" = "#FFA500",
      "Watermelon" = "#800080",
      "Yacon" = "#008080"
    ),
    extra = scale_shape_manual(values = seq(1, length(
      unique(an$timePoint)
    ), 1))
  )
)

cat("\n", "who are those PC3 outliers and why?","\n")

outlier_idx<-which(mod$data$scores[,3]<=-80)
an[outlier_idx,]

cat("\n", "PC3 outliers are from the same person:",unique(an$sourceID[outlier_idx]), "eating ", unique(an$type[outlier_idx]), "collected from",length(outlier_idx),"different timepoints","\n")

PlotLoadSpec(mod,PC = 3)

cat("\n", "From the loadings, it seems that those outliers are from high glucose in urine","\n")


cat("\n", "Since this participant only participated in the Elephant Garlic study it may be wise to remove from the analysis","\n")

# matspec(X_pqn[which(an$sourceID=="fib380"),],ppm,roi = c(3.5,4.0),interactive = F)

```


```{r}
mod1<-PCA(X_pqn[-which(an$sourceID=="fib380"),])

plotScores(
  mod1,
  optns = list(
    color = an$type[-which(an$sourceID=="fib380")],
    shape = an$timePoint[-which(an$sourceID=="fib380")],
    discretePalette = c(
      "Apple" = "#FF0000",
      "Baseline diet only" = "#00FF00",
      "Elephant Garlic" = "#0000FF",
      "Globe Artichoke" = "#FFFF00",
      "Jerusalem Artichoke" = "#FF00FF",
      "Orange" = "#00FFFF",
      "Tomato" = "#FFA500",
      "Watermelon" = "#800080",
      "Yacon" = "#008080"
    ),
    extra = scale_shape_manual(values = seq(1, length(
      unique(an$timePoint)
    ), 1))
  )
)
```



```{r,results='asis'}
roi_list<-list(c(0.5,1.0),c(1.0,1.5),
               c(1.5,2.0),c(2.0,2.5),
               c(2.5,3.0),c(3.0,3.5),
               c(3.5,4.0),c(4.0,4.5),
               c(4.5,5.0),c(5.0,5.5),
               c(5.5,6.0),c(6.0,6.5),
               c(6.5,7.0),c(7.0,7.5),
               c(7.5,8.0),c(8.0,8.5),
               c(8.5,9.0),c(9.0,9.5)
               )
```


## COMPASS
```{r,results='asis',fig.width=6,fig.height=8}

library(gridExtra)
for(i in roi_list){
  model<-PCA(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),which(ppm>=i[1] & ppm<=i[2])],rank = 2)
  
  plotScores(
    model,
    optns = list(
      color = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
      shape = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
      discretePalette = c(
        "Apple" = "#FF0000",
        "Baseline diet only" = "#00FF00",
        "Elephant Garlic" = "#0000FF",
        "Globe Artichoke" = "#FFFF00",
        "Jerusalem Artichoke" = "#FF00FF",
        "Orange" = "#00FFFF",
        "Tomato" = "#FFA500",
        "Watermelon" = "#800080",
        "Yacon" = "#008080"
      ),
      extra = scale_shape_manual(values = seq(1, length(
        unique(an$timePoint)
      ), 1))
    )
  )
  PlotLoadSpec(model,PC = 1)->p1
  PlotLoadSpec(model,PC = 2)->p2
  
  grid.arrange(p1,p2,nrow=2)

  
}
```




## Visual Inspection using specOverlay

soutceID fib 380 (high glucose outlier) were omitted and only ploted at Timepoint T2, T4, T6, and T8

```{r,results='asis',fig.width=10,fig.height=10}
for (i in roi_list) {
  specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(i[1], i[2]),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))->p1
  
  plot(p1)

}

```

### 2,3-butanediol 1.125 ~ 1.150 --> Yacon
```{r,results='asis',fig.width=6,fig.height=10}
specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(1.125,1.150),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```


### lactate 1.3 ~ 1.5 --> Apple at T2
```{r,results='asis',fig.width=6,fig.height=10}

specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(1.31,1.34),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```


### s 1.92 --> Global Artichoke at T2? --> only two samples
```{r,results='asis',fig.width=6,fig.height=10}

specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(1.91,1.93),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```


### s 2.045 --> Elephant Garlic peak at T4
```{r,results='asis',fig.width=6,fig.height=10}

specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(2.0,2.1),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```

### s 3.1 --> Orange 
```{r,results='asis',fig.width=6,fig.height=10}

specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(3.09,3.12),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```

### s 3.3 --> Orange 
```{r,results='asis',fig.width=6,fig.height=10}

specOverlay(X_pqn[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8")),],
              ppm,
              roi = c(3.29,3.31),
              optns = list(facet = an$type[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))],
                           color = an$timePoint[which(an$sourceID!="fib380"&an$timePoint %in% c("T2","T4","T6","T8"))]))
```


